---
title: "p8105_hw2_yt2785"
author: Yijing Tao (UNI yt2785)
date: Oct 2021
output: github_document
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(haven)
library(dplyr)
library(ggplot2)
```


**Problem 1**

Read and clean the Mr. Trash Wheel sheet

```{r warning=FALSE}
trash_wheel_df = 
  read_excel("./data_hw2/Trash_Wheel_Collection_Totals_7_2020_2.xlsx",
             sheet = "Mr. Trash Wheel",
             range =  cell_cols("A:N"), 
             col_names = TRUE, 
             trim_ws = TRUE, 
             skip = 1) %>%
  janitor::clean_names() %>%
  filter(!is.na(dumpster)) %>% 
  mutate(sports_balls = round(sports_balls, 0))

trash_wheel_df
```

Read and clean precipitation data for 2018 and 2019

```{r warning=FALSE}
precipitation_2018_df =
  read_excel("./data_hw2/Trash_Wheel_Collection_Totals_7_2020_2.xlsx",
             sheet = "2018 Precipitation", 
             range = "A2:B14",
             col_names = TRUE, 
             trim_ws = TRUE) %>% 
  janitor::clean_names() %>% 
  filter(!is.na(total)) %>% 
  mutate(year = 2018) %>% 
  select(year, month, total)

precipitation_2019_df =
  read_excel("./data_hw2/Trash_Wheel_Collection_Totals_7_2020_2.xlsx",
             sheet = "2019 Precipitation", 
             range = "A2:B14",
             col_names = TRUE, 
             trim_ws = TRUE) %>% 
  janitor::clean_names() %>% 
  filter(!is.na(total)) %>% 
  mutate(year = 2019) %>% 
  select(year, month, total)

precipitation_df = bind_rows(precipitation_2018_df,precipitation_2019_df) %>% 
  janitor::clean_names() %>% 
  mutate(month = month.name[month])

precipitation_df
```

conclusion data

```{r}
n_row_trash_wheel = nrow(trash_wheel_df)
n_col_trash_wheel = ncol(trash_wheel_df)
trash_wheel_name = colnames(trash_wheel_df)
before_summary_trash_wheel_df = trash_wheel_df %>% 
    janitor::clean_names() %>% 
  select(weight_tons:homes_powered)
summary_trash_wheel_df = knitr::kable(summary(before_summary_trash_wheel_df))
```


```{r}
n_row_precipitation = nrow(precipitation_df)
n_col_precipitation = ncol(precipitation_df)
recipitation_name = colnames(precipitation_df)
before_summary_precipitation_df = precipitation_df %>% 
    janitor::clean_names() %>% 
  select(total)
summary_precipitation_df = knitr::kable(summary(before_summary_precipitation_df))
year_2018_df = filter(precipitation_df, year == "2018")
sum_2018 = sum(pull(year_2018_df, total))
trash_wheel_2019_df = filter(trash_wheel_df, year == 2019)
median_sport_ball_2019 = median(pull(trash_wheel_2019_df, sports_balls))
```
**CONCLUSION 1**
In dataset of Mr. Trash Wheel, there are `r n_row_trash_wheel` observations and `r n_col_trash_wheel` columns in the dataset. The variables includes `r trash_wheel_name` . The summary of the dataset is `r summary_trash_wheel_df`

In precipitation dataframe, there are `r n_row_precipitation` observations and `r n_col_precipitation` columns in the dataset. The variables includes `r recipitation_name` . The summary of the dataset is `r summary_precipitation_df`

Total precipitation in 2018 is `r sum_2018`.
The median number of sports balls in a dumpster in 2019 is `r median_sport_ball_2019`.

**Problem 2**

First, clean the data in pols-month.csv. 

```{r}
pols_month_df = 
  read_csv("./data_hw2/pols-month.csv") %>%
  janitor::clean_names() %>% 
    separate(mon, into = c("year", "month", "day"), sep = "\\-") %>% 
    mutate(month = month.name[as.numeric(month)],
           president = recode(prez_dem, '0' = "gop", '1' = "dem"),
           year = as.numeric(year)) %>% 
  select(-prez_dem, -prez_gop) %>% 
  select(-day)

pols_month_df
```

Second, clean the data in snp.csv using a similar process to the above.

```{r}
snp_df = 
  read_csv("./data_hw2/snp.csv") %>% 
janitor::clean_names() %>% 
   separate(date, into = c("month", "day", "year"), sep = "\\/") %>% 
    mutate(month = month.name[as.numeric(month)]) %>% 
  mutate(year = recode(year, '50' = "1950",'51' = "1951",'52' = "1952",'53' = "1953",'54' = "1954",'55' = "1955",'56' = "1956",'57' = "1957",'58' = "1958",'59' = "1959",'60' = "1960",'61' = "1961",'62' = "1962",'63' = "1963",'64' = "1964",'65' = "1965",'66' = "1966",'67' = "1967",'68' = "1968",'69' = "1969",'70' = "1970",'71' = "1971",'72' = "1972",'73' = "1973",'74' = "1974",'75' = "1975",'76' = "1976",'77' = "1977",'78' = "1978",'79' = "1979",'80' = "1980",'81' = "1981",'82' = "1982",'83' = "1983",'84' = "1984",'85' = "1985",'86' = "1986",'87' = "1987",'88' = "1988",'89' = "1989",'90' = "1990",'91' = "1991",'92' = "1992",'93' = "1993",'94' = "1994",'95' = "1995",'96' = "1996",'97' = "1997",'98' = "1998",'99' = "1999",'00' = "2000",'01' = "2001",'02' = "2002",'03' = "2003",'04' = "2004",'05' = "2005",'06' = "2006",'07' = "2007",'08' = "2008",'09' = "2009",'10' = "2010",'11' = "2011",'12' = "2012",'13' = "2013",'14' = "2014",'15' = "2015")) %>% 
      mutate(year = as.numeric(year)) %>% 
  select(-day) %>% 
  arrange(year, month) %>% 
  relocate(year, month,everything())

snp_df
```

Third, tidy the unemployment data so that it can be merged with the previous datasets.

```{r}
unemployment_df = 
  read_csv("./data_hw2/unemployment.csv") %>% 
  janitor::clean_names() %>% 
pivot_longer(
  jan:dec,
    names_to = "month",
  values_to = "percentage") %>% 
      mutate(year = as.numeric(year)) %>% 
  mutate(month = as.factor(month)) %>% 
  mutate(month = month.name[month])

unemployment_df
```

Join the three datasets above:

```{r}
five_thirty_eight_df = 
  left_join(pols_month_df, snp_df, unemployment_df, by = c("year", "month"))

five_thirty_eight_df
```

conclusion data

```{r}
n_row_pols_month = nrow(pols_month_df)
n_col_pols_month = ncol(pols_month_df)
pols_month_name = colnames(pols_month_df)
pols_month_year_max = max(pull(pols_month_df,year))
pols_month_year_min = min(pull(pols_month_df,year))
before_pols_month_df = pols_month_df %>% 
    janitor::clean_names() %>% 
  select(gov_gop:president)
summary_pols_month_df = knitr::kable(summary(before_pols_month_df))
```


```{r}
n_row_snp = nrow(snp_df)
n_col_snp = ncol(snp_df)
snp_name = colnames(snp_df)
snp_year_max = max(pull(snp_df,year))
snp_year_min = min(pull(snp_df,year))
before_summary_snp_df = snp_df %>% 
    janitor::clean_names() %>% 
  select(close)
summary_snp_df = knitr::kable(summary(before_summary_snp_df))
```


```{r}
n_row_unemployment = nrow(unemployment_df)
n_col_unemployment = ncol(unemployment_df)
unemployment_name = colnames(unemployment_df)
unemployment_year_max = max(pull(unemployment_df,year))
unemployment_year_min = min(pull(unemployment_df,year))
before_unemployment_df = unemployment_df %>% 
    janitor::clean_names() %>% 
  select(percentage)
summary_unemployment_df = knitr::kable(summary(before_unemployment_df))
```

```{r}
n_row_five_thirty_eight = nrow(five_thirty_eight_df)
n_col_five_thirty_eight = ncol(five_thirty_eight_df)
five_thirty_eight_name = colnames(five_thirty_eight_df)
five_thirty_eight_year_max = max(pull(five_thirty_eight_df,year))
five_thirty_eight_year_min = min(pull(five_thirty_eight_df,year))
before_five_thirty_eight_df = five_thirty_eight_df %>% 
    janitor::clean_names() %>% 
  select(gov_gop:close)
summary_five_thirty_eight_df = knitr::kable(summary(before_five_thirty_eight_df))
```
**CONCLUSION 2**
In the pol month dataset, there are `r n_row_pols_month` rows and `r n_col_pols_month` columns in the dataset. The data cover from `r pols_month_year_min`  to `r `pols_month_year_max`.The variables includes `r pols_month_name` . The summary of the dataset is `r summary_pols_month_df`

In the snp dataset, there are `r n_row_snp` rows and `r n_col_snp` columns in the dataset. The data cover from `r snp_year_min`  to `r snp_year_max`.The variables includes `r snp_name` . The summary of the dataset is `r summary_snp_df`

In the unemployment dataset, there are `r n_row_unemployment` rows and `r n_col_unemployment` columns in the dataset. The data cover from `r unemployment_year_min`  to `r unemployment_year_max` .The variables includes `r unemployment_name`. The summary of the dataset is `r summary_unemployment_df`

In the joint dataset, there are `r n_row_five_thirty_eight` rows and `r n_col_five_thirty_eight` columns in the dataset. The data cover from  `r five_thirty_eight_year_min` to `r five_thirty_eight_year_max` . The variables includes `r five_thirty_eight_name` . The summary of the dataset is `r summary_five_thirty_eight_df`


**Problem 3**

Load and tidy the data.

```{r}
popular_baby_names_df = 
  read_csv("./data_hw2/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(ethnicity = recode(ethnicity, 'ASIAN AND PACI' = "ASIAN AND PACIFIC ISLANDER", 'BLACK NON HISP' = "BLACK NON HISPANIC", 'WHITE NON HISP' = "WHITE NON HISPANIC")) %>% 
  distinct(year_of_birth,gender,ethnicity,childs_first_name,count,rank, .keep_all = TRUE)

popular_baby_names_df
```

Produce a well-structured, reader-friendly table showing the rank in popularity of the name `Olivia` as a female baby name over time

```{r}
olivia_df = filter(popular_baby_names_df, childs_first_name == "Olivia", gender == "FEMALE") %>% 
  select(year_of_birth, ethnicity, rank) %>% 
 pivot_wider(
    names_from = "year_of_birth",#name from another column
    values_from = "rank"#values from another column
  ) 
knitr::kable(olivia_df)
```

 Produce a similar table showing the most popular name among male children over time.

```{r}
popular_male_name_df = filter(popular_baby_names_df,gender == "MALE", rank == 1) %>% 
  select(year_of_birth, ethnicity, childs_first_name) %>% 
 pivot_wider(
    names_from = "year_of_birth",#name from another column
    values_from = "childs_first_name"#values from another column
  ) 
knitr::kable(popular_male_name_df)
```

Finally, for male, white non-hispanic children born in 2016, produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis).

```{r}
male_white_2016_df = filter(popular_baby_names_df,gender == "MALE", ethnicity == "WHITE NON HISPANIC", year_of_birth == 2016) %>% 
  select(childs_first_name, rank, count)

ggplot(male_white_2016_df, aes(x = rank, y = count)) + geom_point()
```
